---
title: "p8105_hw3_lg3156"
author: "Liner Ge"
date: "2020/10/7"
output: github_document
---

```{r setup}
library(tidyverse)
library(p8105.datasets)
library(readxl)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes.

### How many aisles, and which are most items from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are 134 aisles and most items from fresh vegetables.

### Make a plot

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

### Make a table

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
   group_by(aisle) %>% 
   count(product_name) %>% 
   mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

### Apples vs ice cream

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream") ) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour=mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from= mean_hour
  )
```

# Problem 2

### Load, tidy, and otherwise wrangle the data

```{r}
activity_df=
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to= "minute",
    names_prefix = "activity_",
    values_to ="counts") %>% 
  mutate(
    weekday_weekend = case_when(
      day %in% c("Friday", "Monday", "Thursday", "Tuesday", "Wednesday") ~ "weekday",
      day %in% c("Saturday", "Sunday") ~ "weekend",
    )
  ) %>% 
  mutate(
    week = as.factor(week),
    day_id= as.numeric(day_id),
    minute= as.numeric(minute),
    weekday_weekend = as.factor(weekday_weekend),
    day = as.factor(day),
    counts = as.numeric(counts)
  ) %>% 
  mutate(
     day = forcats::fct_relevel(day,"Monday","Tuesday","Wednesday","Thursday","Friday", "Saturday","Sunday")
  ) %>% 
  group_by(week) %>% 
  arrange(day,.by_group = TRUE) %>% 
  relocate (day_id, week, weekday_weekend)
```

This dataset includes variables --  day_id, week, weekday_weekend, day, minute, and counts. This dateset has `r ncol(activity_df) ` columns and contains `r nrow(activity_df)` observations. 

### Traditional analyses of accelerometer data

```{r}
activity_df %>% 
  group_by(week, day) %>% 
  summarize(
    day_activity_counts = sum (counts)
  ) %>% 
  pivot_wider(
    names_from = day,
    values_from = day_activity_counts
  )%>% 
  knitr::kable(digits = 2)
```

From this table, I can notice that on Saturday in last two week, the man had the least action. It seems like the man had less action on weekend than on weekdays. However, the counts of action is fluctuant among 5 weeks. Thus, I cannot find other trend apparently. 

### A plot showing the 24-hour activity

```{r}
activity_df %>% 
  mutate(
    hour = (minute-1)%/%60+1
  ) %>% 
  group_by(day_id, week, day, hour) %>% 
  summarize(
    hour_activity_counts = sum (counts)
  ) %>% 
  ggplot( aes(x = hour, y = hour_activity_counts, color = day, group = day_id))+
  geom_smooth(aes(linetype = week),se = FALSE)+
  scale_y_continuous(trans ="sqrt",
                     name = "Activity counts in an hour")+
  scale_x_continuous(name = "Hour")+
  labs(
    title = "The 24-hour activity counts in 35 days",
    caption = "Data from accelerometer data")
```

In most of the days, the 63 year-old man's main activity time is between 10 am and 8 pm. There is no significant different between weekday and weekend. 